---
title: "Test modifying epinowcast"
output: html_document
date: "2025-01-13"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(epinowcast)
library(data.table)
library(ggplot2)
```

The goal of this Rmd is to practice modifying the stan model in epinowcast and
to get familiar with the model specification/input formats, as we might
use that to make a simple baseline nowcasting model

We will use the same data from the epinowcast vignette (German hospital 
admissions by reference and report date)

```{r}
nat_germany_hosp <-
  germany_covid19_hosp[location == "DE"][age_group == "00+"] |>
  enw_filter_report_dates(latest_date = "2021-10-01")

retro_nat_germany <- nat_germany_hosp |>
  enw_filter_report_dates(remove_days = 40) |>
  enw_filter_reference_dates(include_days = 40)
retro_nat_germany

latest_germany_hosp <- nat_germany_hosp |>
  enw_latest_data() |>
  enw_filter_reference_dates(remove_days = 40, include_days = 40)
head(latest_germany_hosp, n = 10)
```
Convert into "reporting triangle" format  and specify the model. 
Things to specify:
- number of days of data to use for the reference and report modules
- maximum delay to consider 
- grouping (age group, location, both) of observations

```{r}
pobs <- enw_preprocess_data(retro_nat_germany, max_delay = 40)
```

Specify the process model: Assumes exponential growth rate model with a 
random effect on day (equivalent to a geometric random walk on expected 
counts by reference date)
```{r}
expectation_module <- enw_expectation(
  ~ 0 + (1 | day), data = pobs
)
```
Reporting model by reference date: Assumes reporting delay is lognormally
distributed and static over time and groups. 
```{r}
reference_module <- enw_reference(~1, distribution = "lognormal", data = pobs)
```
Could also have modeled the reporting delay non-parameterically using a hazard
model, or some combination of parametric and non-parameteric.
Reporting effects by report date: model the effects on reporting processes using
a hazard formulation using enw_report().
```{r}
report_module <- enw_report(~ (1 | day_of_week), data = pobs)
```
Precompile the model
```{r}
model <- enw_model(threads = TRUE)
```
Fitting the model:
```{r}
options(mc.cores = 2)
nowcast <- epinowcast(data = pobs,
  expectation = expectation_module,
  reference = reference_module,
  report = report_module,
  fit = enw_fit_opts(
    save_warmup = FALSE, pp = TRUE,
    chains = 2, threads_per_chain = 2,
    iter_sampling = 500, iter_warmup = 500,
    show_messages = FALSE
  ),
  model = model
)
```
